---
title: "Answers"
date: "`r Sys.Date()`"
---

## PART ONE: Loading Data & Data Wrangling

All of the following questions are based on the sample of restaurants represented in the dataset which is accurate as of September 2, 2016.

# Load Packages
```{r}
library(tidyverse)
```

# Load Data

```{r}
brodhead <- read_csv("data/brodheadCenter.csv")
```

# Exercises: Data Wrangling

1. Which restaurant has the lowest cost item and what is the item?

```{r}
brodhead %>% 
  arrange(cost) %>% 
  select(name, type, itemName, cost)
```


2. Which restaurant has the most expensive item(s)?  What are those item(s)?

```{r}
brodhead %>% 
  arrange(desc(cost)) %>% 
  select(name, type, itemName, cost)
```

##### Alternative Answer

```{r include=FALSE}
answer <- brodhead %>% 
  arrange(desc(cost)) %>% 
  filter(type == "restaurant") %>% 
  select(name, type, itemName, cost) %>% 
  head(1) 

answer2 <- brodhead %>% 
  select(name, type, itemName, cost) %>% 
  filter(type == "restaurant",
         cost == max(cost)) 
```


The **`r answer$name`** *`r answer$type`* has the most expensive item(s) at $`r answer$cost`.  Those item(s) are `r answer$itemName`.

The **`r answer2$name[1]`** *`r answer2$type[1]`* has the most expensive item(s) @ $`r answer2$cost[1]`.  Those item(s) are `r answer2$itemName`.

3. At the Brodhead Center, how many of the entrees (found in the `menuType` variable) cost eight dollars?

```{r}
brodhead %>% 
  filter(cost == 8, menuType == "entree") %>% 
  select(name, menuType, itemName, cost)
```


4. The head of Duke dining is considering reducing prices at the Brodhead Center. Using what we've learned in class, write a line of code that will calculate a new variable (`halfPrice`) that contains items at half price.

```{r}
brodhead %>% 
  mutate(halfPrice = cost / 2) %>% 
  select(name, itemName, cost, halfPrice)
```


5. How many entrees are in the dataset (`menuType` variable)? How many desserts 
```{r}
brodhead %>% 
  count(menuType)
```


# PART TWO:  Data Structures & Vector Types

Use the Brodhead Center prebuilt a dataset that has information about menus from three restaurants in Duke's dining hall.  

You should have created a new R Notebook, and each notebook should be self contained.  This means you should load the packages and data again, even though you did that in the previous exercise.

1. Using what we've you've seen in class, if `brodheadCenter.csv` is a csv file, how would you load the file into a new object called `brodhead`?

```{r}
library(tidyverse)
brodhead <- read_csv("data/brodheadCenter.csv")
# or 
# brodhead <- read_csv("https://raw.githubusercontent.com/data-and-visualization/Intro2R/master/data/brodheadCenter.csv")
```

2. What data structure is `brodhead`?

```{r}
class(brodhead)
```


3. Take a look at the structure of the `brodhead` object.
    - How many observations (rows) are there?
    - How many variables (columns) are there?
    - How many of the variables are numeric data?
    
```{r}
glimpse(brodhead)
```
